A Random Choice SPH Scheme with Adaptive Viscosity

### 2019-10-28 ebp-JCPManuscript.tex
update tex file to use the new style file

\documentclass[review]{elsarticle}
\usepackage{lineno,hyperref}
% package added by zhixuan, not sure whether allowed by JCP or not!
\usepackage{amsmath}
\usepackage{float}

\modulolinenumbers[5]
\journal{Journal of Computational Physics}
%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%
%% Numbered
\bibliographystyle{model1-num-names}
%% Numbered without titles
%\bibliographystyle{model1a-num-names}
%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}
%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}
%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}
%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{frontmatter}
\title{A Random Choice SPH Scheme with Adaptive Viscosity}
%\tnotetext[mytitlenote]{Can add  title note here}
%% Group authors per affiliation:
\author{Zhixuan Cao}
\author{Abani Patra \fnref{mycorrespondingauthor}}
\address{Department of Mechanical and Aerospace Engineering, The State University of New York at Buffalo, Amherst, NY, 14260-4200, USA}
\author{E. Bruce Pitman}
\address{Department of Materials Design and Innovation, The State University of New York at Buffalo, Amherst, NY, 14260-4200, USA}
%\fntext[myfootnote]{Since 1880.}
%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]%{Elsevier Inc}
%\ead[url]{www.elsevier.com}
%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
\fntext[mycorrespondingauthor]{abani@buffalo.edu}
%\cortext[mycorrespondingauthor]{abani@buffalo.edu}
%\ead{abani@buffalo.edu}
%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

@@

\linenumbers
\section{Introduction}
Smoothed particle hydrodynamics (SPH) \citep{gingold1977smoothed,lucy1977numerical} is a mesh-free particle method based on a Lagrangian formulation of the governing equations of motion.
SPH has been widely applied to different areas in engineering and science. As should be expected, SPH possesses certain advantages and certain disadvantages relative to finite difference and finite element solvers. Principally, as a Lagrangian solver, SPH conserves mass exactly. 
In addition, because of its formulation SPH automatically concentrates numerical effort in regions of high density and adapts to arbitrary geometries. On the other hand, 
SPH has difficulty capturing shock waves and contact discontinuities. To overcome numerical oscillations at discontinuities, classical SPH introduces an artificial viscosity term. As it is most often used, when two SPH particles are approaching, a von Neumann type viscosity is introduced
\citep{monaghan1983shock, monaghan1997sph,klapp2012strong}. 
This kind of viscosity, however, usually provides more dissipation than is truly needed at a discontinuity, and the excess dissipation ultimately corrupts the entire computation. Various methods have been introduced to tune the artificial viscosity, turning it on and off for particular wave features. However, tuning switches cannot properly accommodate every kind of wave that might arise, and spurious effects and unphysical damping persist (see, for example, \cite{borgani2012hydrodynamic,flebbe1994smoothed,morris1997switch, dolag2005turbulent,cullen2010inviscid,sigalotti2008adaptive,puri2014comparison}).

@@

In contrast, Glimm\rq{}s approach solves a local Riemann problem exactly within each cell,
and samples (randomly or uniformly) the solution states. Consider again the linear advection equation \ref{linear}, with a discontinuity in cell $j$:
\begin{equation}
u^n = \left\{ \begin{array}{11}
			u_L ~~~ x< (j+\xi)\Delta x \\
 			u_R ~~~ x> (j+\xi)\Delta x
 			\end{array}
	  \right. 			
\end{equation}
where $0<\xi<1$ locates the discontinuity within cell $j$.

At time $n+1$, the discontinuity is either propagated or not, depending on the sampling
point:
\begin{equation}
u^{n+1} = \left\{\begin{array}{11}
		  u_L ~~~ x< (j+\tilde \xi)\Delta x + a \Delta t \\
		  u_R ~~~ x> (j+\tilde \xi)\Delta x + a \Delta t
		  \end{array}
		  \right.
\end{equation}
where $\tilde \xi$ denotes the sample point.
In either case, the solution remains discontinuous at time $n+1$ and the 
discontinuity propagates at speed $a$ on average.

@@

Adaptive smoothing length is now an essential component of SPH method. It naturely results to nonuniform smoothing length in space. Variation range of smoothing length would be especially larger for compressible flow. Fortunately, there are several factors counteracting against the numerical oscillation:
\begin{itemize}
\item Numerical dissipation decays such numerical oscillation preventing accumulation of oscillation magnitude.
\item For incompressible flow or weakly compressible flow, smoothing length only varies in a small range.
\item Well developed adaptive smoothing length algorithm usually guarantees more ``smoothing" change of smoothing length in space avoiding sharp change of smoothing length in space.
\end{itemize}
So we do not think that the numerical oscillation is a challenging issue for SPH. The major motivation of the appendix is to show that 
\begin{itemize}
\item The numerical oscillation does exist commonly in SPH schemes. It is not something only for RSPH.
\item Dissipation decays the oscillation, so numerical oscillation can serve as an indicator of dissipation.
\end{itemize}
\section*{References}

\bibliography{Reference}

>>>>

%% 
%% Copyright 2007-2019 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references

\documentclass[preprint,12pt,authoryear]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times,authoryear]{elsarticle}
%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,3p,times,authoryear]{elsarticle}
%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,5p,times,authoryear]{elsarticle}
%% \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
% package added by zhixuan
\usepackage{amsmath}
\usepackage{float}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

\journal{Computer Methods in Applied Mechanics and Engineering}
\begin{document}
\begin{frontmatter}
\title{A Random Choice SPH Scheme with Adaptive Viscosity}
%\tnotetext[mytitlenote]{Can add  title note here}
%% Group authors per affiliation:
\author{Zhixuan Cao}
\ead{zhixuan.cao@ansys.com}
\address{Fluids Business Unit, Ansys Inc., Lebanon, NH, 03766, USA}
\author{Abani Patra \fnref{mycorrespondingauthor}}
\ead{abani.patra@tufts.edu}
\address{Department of Computer Science and Department of Mathematics, Tufts University, Medford, MA, 02153, USA}
\author{E. Bruce Pitman}
\ead{pitman@buffalo.edu}
\address{Department of Materials Design and Innovation, The State University of New York at Buffalo, Amherst, NY, 14260, USA}
%\fntext[myfootnote]{Since 1880.}
%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]%{Elsevier Inc}
%\ead[url]{www.elsevier.com}
%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
%\cortext[mycorrespondingauthor]{abani@buffalo.edu}
%\ead{abani@buffalo.edu}
%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

@@

\section{Introduction}
Smoothed particle hydrodynamics (SPH) \citep{gingold1977smoothed,lucy1977numerical} is a mesh-free particle method based on a Lagrangian formulation of the governing equations of motion.
SPH has been widely applied to different areas in engineering and science. As should be expected, SPH possesses certain advantages and certain disadvantages relative to finite difference and finite element solvers. Principally, as a Lagrangian solver, SPH conserves mass exactly. 
In addition, because of its formulation SPH automatically concentrates numerical effort in regions of high density and adapts to arbitrary geometries. On the other hand, 
SPH has difficulty capturing shock waves and contact discontinuities. To overcome numerical oscillations at discontinuities, classical SPH introduces an artificial viscosity term. As it is most often used, when two SPH particles are approaching, a von Neumann type viscosity is introduced
\citep{monaghan1983shock, monaghan1997sph,klapp2012strong}. 
This kind of viscosity, however, usually provides more dissipation than is truly needed at a discontinuity, and the excess dissipation ultimately corrupts the entire computation. Various methods have been introduced to tune the artificial viscosity, turning it on and off for particular wave features. However, tuning switches cannot properly accommodate every kind of wave that might arise, and spurious effects and unphysical damping persist (see, for example, \cite{borgani2012hydrodynamic,flebbe1994smoothed,morris1997switch, dolag2005turbulent,cullen2010inviscid,sigalotti2008adaptive,puri2014comparison}).

@@

In contrast, Glimm\rq{}s approach solves a local Riemann problem exactly within each cell,
and samples (randomly or uniformly) the solution states. Consider again the linear advection equation \ref{linear}, with a discontinuity in cell $j$:
\begin{equation}
u^n = \begin{cases}
	u_L ~~~ x< (j+\xi)\Delta x \\
 	u_R ~~~ x> (j+\xi)\Delta x
	\end{cases}			
\end{equation}
where $0<\xi<1$ locates the discontinuity within cell $j$.

At time $n+1$, the discontinuity is either propagated or not, depending on the sampling
point:
\begin{equation}
u^{n+1} = \begin{cases}
		  u_L ~~~ x< (j+\tilde \xi)\Delta x + a \Delta t \\
		  u_R ~~~ x> (j+\tilde \xi)\Delta x + a \Delta t
	        \end{cases}
\end{equation}
where $\tilde \xi$ denotes the sample point.
In either case, the solution remains discontinuous at time $n+1$ and the 
discontinuity propagates at speed $a$ on average.

@@

Adaptive smoothing length is now an essential component of SPH method. It naturely results to nonuniform smoothing length in space. Variation range of smoothing length would be especially larger for compressible flow. Fortunately, there are several factors counteracting against the numerical oscillation:
\begin{itemize}
\item Numerical dissipation decays such numerical oscillation preventing accumulation of oscillation magnitude.
\item For incompressible flow or weakly compressible flow, smoothing length only varies in a small range.
\item Well developed adaptive smoothing length algorithm usually guarantees more ``smoothing" change of smoothing length in space avoiding sharp change of smoothing length in space.
\end{itemize}
So we do not think that the numerical oscillation is a challenging issue for SPH. The major motivation of the appendix is to show that 
\begin{itemize}
\item The numerical oscillation does exist commonly in SPH schemes. It is not something only for RSPH.
\item Dissipation decays the oscillation, so numerical oscillation can serve as an indicator of dissipation.
\end{itemize}
\section*{References}
 \bibliographystyle{elsarticle-num} 
\bibliography{Reference}
